---
title: "CBNplot"
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
vignette: >
  %\VignetteIndexEntry{CBNplot}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# CBNplot: Bayesian network plot for clusterProfiler results


## Introduction

The R package to infer and plot Bayesian networks. The network are inferred from expression data based on [clusterProfiler](https://github.com/YuLab-SMU/clusterProfiler) or ReactomePA results. It makes use of libraries including [clusterProfiler](https://github.com/YuLab-SMU/clusterProfiler), [ReactomePA](https://github.com/YuLab-SMU/ReactomePA), [bnlearn](https://www.bnlearn.com/), [graphite](https://bioconductor.org/packages/release/bioc/html/graphite.html) and [depmap](https://bioconductor.org/packages/release/data/experiment/html/depmap.html). In this vignette, the description of functions and several use cases are depicted using [GSE133624](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE133624), which contains RNA-Seq data of bladder cancer. The more detail can be found on the book ([https://noriakis.github.io/CBNplot/](https://noriakis.github.io/CBNplot/)).

## Installation
```{r, eval=FALSE}
BiocManager::install("CBNplot")
```

## Usage
### The preprocessing and DEG identification of GSE133624
```{r deg, include=TRUE, echo=TRUE, message=FALSE, cache=FALSE, warning=FALSE, comment=FALSE, fig.height = 10, fig.width = 10}
library(CBNplot)
library(DESeq2)
library(org.Hs.eg.db)
library(GEOquery)
## Load dataset and make metadata
filePaths <- getGEOSuppFiles("GSE133624")
counts = read.table(rownames(filePaths)[1], header=1, row.names=1)
meta = sapply(colnames(counts), function (x) substring(x,1,1))
meta = data.frame(meta)
colnames(meta) = c("Condition")

dds <- DESeqDataSetFromMatrix(countData = counts,
                              colData = meta,
                              design= ~ Condition)
## Prefiltering
filt <- rowSums(counts(dds) < 10) > dim(meta)[1]*0.9
dds <- dds[!filt,]

## Perform DESeq2()
dds = DESeq(dds)
res = results(dds, pAdjustMethod = "bonferroni")

## apply variance stabilizing transformation
v = vst(dds, blind=FALSE)
vsted = assay(v)

## Define the input genes, and use clusterProfiler::bitr to convert the ID.
sig = subset(res, padj<0.05)
cand.entrez = clusterProfiler::bitr(rownames(sig),
  fromType="ENSEMBL", toType="ENTREZID", OrgDb=org.Hs.eg.db)$ENTREZID

## Perform enrichment analysis
pway = ReactomePA::enrichPathway(gene = cand.entrez)
pway = clusterProfiler::setReadable(pway, org.Hs.eg.db)

## Define including samples
incSample = rownames(subset(meta, Condition=="T"))
```
### The use of CBNplot


## bngeneplot

Then use CBNplot. Basically, you need to supply the enrichment analysis result, normalized expression value and samples to be included. For `bngeneplot`, the pathway number in the `result` slot of enrichment analysis results must be given.

```{r usecase, include=TRUE, echo=TRUE, message=FALSE, cache=FALSE, warning=FALSE, comment=FALSE, fig.height = 10, fig.width = 10}
bngeneplot(results = pway,exp = vsted,
  expSample = incSample, pathNum = 15)
```

Data frame of raw values used in the inference, data frame containing strength and direction, averaged network, and plot can be obtained by specifying `returnNet=TRUE`
```{r usecase2, include=TRUE, echo=TRUE, message=FALSE, cache=FALSE, warning=FALSE, comment=FALSE, fig.height = 10, fig.width = 10}
ret <- bngeneplot(results = pway,exp = vsted,
  expSample = incSample, pathNum = 15, returnNet=TRUE)
ret$str |> head()
```

## bnpathplot

The relationship between pathways can be drawn by `bnpathplot`. The number to be included in the inference can be specified by `nCategory`.

```{r usecase3, include=TRUE, echo=TRUE, message=FALSE, cache=FALSE, warning=FALSE, comment=FALSE, fig.height = 10, fig.width = 10}
bnpathplot(results = pway,exp = vsted,
  expSample = incSample, nCategory=10, shadowText = TRUE)
```

## bngeneplotCustom and bnpathplotCustom

`bngeneplotCustom` and `bnpathplotCustom` can be used to customize visualization with more flexibility, like highlighting the nodes and edges of interest by `glowEdgeNum` and `hub`.

```{r usecase4, include=TRUE, echo=TRUE, message=FALSE, cache=FALSE, warning=FALSE, comment=FALSE, fig.height = 10, fig.width = 10}
bnpathplotCustom(results = pway, exp = vsted, expSample = incSample,
  fontFamily="serif", glowEdgeNum=3, hub=3)
bngeneplotCustom(results = pway, exp = vsted, expSample = incSample,
  pathNum=15, fontFamily="sans", glowEdgeNum=NULL, hub=3)
```

The detailed usage for the package, like including covariates to the plot and probabilistic reasoning is available in the package documentation ([https://noriakis.github.io/CBNplot/](https://noriakis.github.io/CBNplot/)).

```{r}
sessionInfo()
```



